---
# main script to identify the FastIIX, FastIIB, MuSC, FAPs major cell-types

title: "scRNA analysis"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true # table of content true
    toc_depth: 1
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: false  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
---

# Load cell-ranger feature_by_cell matrices 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(stringr)
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2) # version should be >=3.5.0 from CRAN
library(openxlsx)
library(harmony)
library(SingleR)
library(reticulate)
library(MuDataSeurat)

# Directories
proj_path <- "/ix/djishnu/Zarifeh/ML_MM"
data_path <- file.path(proj_path, "Aditi/")
sc_sample_paths <- paste0(data_path, rep(c("aging_F7", "aging_F15", "aging_M4", "aging_M11")))
count_paths <- file.path(sc_sample_paths, "outs/filtered_feature_bc_matrix") # input to seurat from cell-ranger
out_path <- "/ix/djishnu/Akanksha/snRNA_TA_muscle_analysis/Seurat_analysis_outputs"
sample_names <- c("F7_KO", "F15_WT", "M4_KO", "M11_WT")
```

############################################## CELL EMBEDDINGS #########################################################

# Load the QC filtered seurat object with cell embeddings 

``` {r Reading seurat object}
# reads the QC pre-processed seurat obj, which contains one layer for all sample cells (merged), and 10 cell-type labels
cluster_seurat <- readRDS("/ix/djishnu/Zarifeh/ML_MM/Aditi/result/Aging/Agingcluster_seurat.rds")
```

# Load genes/biomarkers of interest 

```{r markers from Atlas}
markers <- c(
  "Ano4", "Prkar1a", "Ufsp1", "Col13a1", # NMJ
  "Tnxb", "Plxdc2", "Egfr", "Gsn", # FAP
  "Abi3bp", "Ptprd", "Nlan1", # Tendon
  "Col22a1", "Adamts20", "Anqpt1", "App", # MTJ
  "Trpc3", "Cacna1c", "Opcml", "Slc35f1", # Pericyte
  "Hs6st3", "Chodl", "Pax7", "Tenm4", # Satellite cell
  "Flt1", "Mecom", "Cyvr1", "Dach1", "Pecam1", # EC
  "F13a1", "Mctp1", "Rbpi", "Slc9a9", # Macrophages
  "Actn3", "Sorbs2", "Prkaq3", "Xpr1", # FastIIB
  "Myh1", "Kcnn2", "Acss2", "Fhl1" # FastIIX
)
```

# Plot marker genes for all identified clusters

```{r violin plot cluster_biomarkers}
VlnPlot(cluster_seurat,
  features = markers, assay = "RNA",
  group.by = "seurat_clusters",
  stack = TRUE, flip = TRUE
)
# ggsave(file.path(out_path, "/plots/Vlnplot_markers.png"))
```

# Augment meta_data for this object

```{r meta data inspection}
metadata <- cluster_seurat@meta.data
sample_names <- c("F7_KO", "F15_WT", "M4_KO", "M11_WT")
# remove some redundant columns
metadata <- metadata %>% select(-sample, -cells, -RNA_snn_res.0.8)
# add celltype_label column
metadata$cell_type <- Idents(cluster_seurat)
# assign the condition based on if F7,M4 then KO or else WT
metadata$condition <- ifelse(metadata$orig.ident %in% c("F7_KO", "M4_KO"), "KO", "WT")
# assign the gender based on if F7,F15 then F or else M
metadata$gender <- ifelse(metadata$orig.ident %in% c("F7_KO", "F15_WT"), "F", "M")
cluster_seurat@meta.data <- metadata
print(head(metadata))
```

# UMAPs : cluster labels based on major cell-types // samples based on conditions and batches

```{r plot all_sample umaps with specific cluster labels}
# check the cluster labels and number of cells per cluster
current.cluster.ids <- levels((Idents(cluster_seurat)))
print(current.cluster.ids)
cell_counts_per_cluster <- table(Idents(cluster_seurat))
print(cell_counts_per_cluster)

# print dimensions of umap reductions (cell embeddings and gene loadings)
umap_embeddings <- cluster_seurat@reductions$umap@cell.embeddings
print(dim(umap_embeddings))

# plotting umaps
umap_cell_type <- DimPlot(cluster_seurat, reduction = "umap", label = TRUE, pt.size = 1)
# ggsave(file.path(out_path, "/plots/all_cells_celltype_labels.png"), plot = umap_cell_type, width = 10.1, height = 10.1)
umap_sample <- DimPlot(cluster_seurat, reduction = "umap", label = TRUE, group.by = "orig.ident", pt.size = 1)
# ggsave(file.path(out_path, "/plots/all_cells_sample_labels.png"), plot = umap_sample, width = 10.1, height = 10.1)
umap_condition <- DimPlot(cluster_seurat, reduction = "umap", label = TRUE, group.by = "condition", pt.size = 1)
# ggsave(file.path(out_path, "/plots/all_cells_condition_labels.png"), plot = umap_condition, width = 10.1, height = 10.1)
```

```{r split the whole seurat object into condition specific objects}
# split the object based on condition metadata
condition_obj_list <- SplitObject(cluster_seurat, split.by = "condition")
condition_obj_list$KO
print(head(condition_obj_list$KO@meta.data))
condition_obj_list$WT
print(head(condition_obj_list$WT@meta.data))
```

```{r plot conditon specific umaps}
# plot umaps based on cell-type labels
umap_KO <- DimPlot(condition_obj_list$KO, reduction = "umap", label = TRUE, pt.size = 1)
# ggsave(file.path(out_path, "/plots/KO_celltype_labels.png"), plot = umap_KO, width = 10.1, height = 10.1)
umap_WT <- DimPlot(condition_obj_list$WT, reduction = "umap", label = TRUE, pt.size = 1)
# ggsave(file.path(out_path, "/plots/WT_celltype_labels.png"), plot = umap_WT, width = 10.1, height = 10.1)
```

```{r save rds}
saveRDS(condition_obj_list$KO, file = file.path(out_path, "/objects/Seurat_KO_major_celltype.rds"))
saveRDS(condition_obj_list$WT, file = file.path(out_path, "/objects/Seurat_WT_major_celltype.rds"))
```

###### Get anndata objects for each condition

```{r save anndata .h5ad}
# read pre-made rds objects
KO_seurat <- readRDS(file.path(out_path, "/objects/Seurat_KO_major_celltype.rds"))
WT_seurat <- readRDS(file.path(out_path, "/objects/Seurat_WT_major_celltype.rds"))
# print(head(KO_seurat@meta.data))

# add cell type and counts to meta data column from idents
# convert seurat objects to anndata objects
MuDataSeurat::WriteH5AD(KO_seurat, file.path(out_path, "/objects/KO_major_celltype.h5ad"), assay = "RNA")
MuDataSeurat::WriteH5AD(WT_seurat, file.path(out_path, "/objects/WT_major_celltype.h5ad"), assay = "RNA")
```

# Biological batch effect between M/F samples per condition

```{r plot batch effect in condition umaps}
# plot umaps based on cell-type labels
umap_KO_sample <- DimPlot(condition_obj_list$KO, reduction = "umap", label = TRUE, group.by = "orig.ident", pt.size = 1)
# ggsave(file.path(out_path, "/plots/KO_sample_labels.png"), plot = umap_KO_sample, width = 10.1, height = 10.1)
umap_WT_sample <- DimPlot(condition_obj_list$WT, reduction = "umap", label = TRUE, group.by = "orig.ident", pt.size = 1)
# ggsave(file.path(out_path, "/plots/WT_sample_labels.png"), plot = umap_WT_sample, width = 10.1, height = 10.1)
```

```{r run harmony integration/batch-effect reduction}
# old reductions
# print(head(condition_obj_list$KO@reductions$harmony@cell.embeddings))
# print(head(condition_obj_list$WT@reductions$harmony@cell.embeddings))

# run harmony to integrate M/F samples per seurat object (for both WT and KO) and add to its reductions slot
KO_batch_integrated_obj <- RunHarmony(condition_obj_list$KO, group.by.vars = "orig.ident") # orig.ident stores the sample name/batch
WT_batch_integrated_obj <- RunHarmony(condition_obj_list$WT, group.by.vars = "orig.ident")
# run umap post this for better viz

# new reductions
print(head(KO_batch_integrated_obj@reductions$harmony@cell.embeddings))
print(head(WT_batch_integrated_obj@reductions$harmony@cell.embeddings))
```

```{r plot harmony integration/batch-effect reduction}
# run UMAP to transform the coordinates of harmony for plotting
nDims <- 18
KO_batch_integrated_obj <- RunUMAP(KO_batch_integrated_obj, dims = 1:nDims)
WT_batch_integrated_obj <- RunUMAP(WT_batch_integrated_obj, dims = 1:nDims)

# plot umaps based on cell-type labels
KO_harmony <- DimPlot(KO_batch_integrated_obj, reduction = "umap", label = TRUE, group.by = "orig.ident", pt.size = 1)
ggsave(file.path(out_path, "/plots/KO_harmony.png"), plot = KO_harmony, width = 10.1, height = 10.1)
WT_harmony <- DimPlot(WT_batch_integrated_obj, reduction = "umap", label = TRUE, group.by = "orig.ident", pt.size = 1)
ggsave(file.path(out_path, "/plots/WT_harmony.png"), plot = WT_harmony, width = 10.1, height = 10.1)
```

########################################### DEG analysis ######################################################
# Identify DEGs between KO experiment and WT controls per cell-type specific cluster; save as .csv

```{r updating orig.ident with sample_condition_celltype}
library(EnhancedVolcano)

levels(cluster_seurat)
seurat_obj <- cluster_seurat
matrix <- seurat_obj@meta.data
matrix$orig.ident <- paste(matrix$orig.ident, matrix$cell_type, sep = "_")
head(matrix)
Idents(seurat_obj) <- matrix$orig.ident
```

```{r FastIIB all DEGs}
FastIIB.de.markers <- FindMarkers(seurat_obj, ident.1 = c("F7_KO_Fast IIB", "M4_KO_Fast IIB"), ident.2 = c("F15_WT_Fast IIB", "M11_WT_Fast IIB"), verbose = TRUE)
head(FastIIB.de.markers)
# write.csv(FastIIB.de.markers, file.path(out_path, "/tables/FastIIB_unfiltered_KO_DEGs.csv"))
df <- FastIIB.de.markers

# plot DEGs
significance_filter <- 0.01
FC_threshold <- 1
Vol_DEG <- EnhancedVolcano(df,
  rownames(df),
  x = "avg_log2FC",
  y = "p_val_adj",
  pCutoff = significance_filter,
  FCcutoff = FC_threshold,
  pointSize = 2,
  drawConnectors = TRUE,
  title = "Fast IIB: All DEGs in KO vs. WT controls",
)
# ggsave(file.path(out_path, "/plots/FastIIB_KO_DEGs.png"), plot = Vol_DEG, width = 10.1, height = 10.1)
```

```{r FastIIB subset up and down regulated DEGs}
# subset up and down regulated DEGs
FastIIB_upregulated_KO_DEGs <- subset(FastIIB.de.markers, p_val_adj < 0.01 & avg_log2FC > 1)
# cat(rownames(FastIIB_upregulated_KO_DEGs), sep = "\n")
# write.csv(FastIIB_upregulated_KO_DEGs, file.path(out_path, "/tables/FastIIB_upregulated_KO_DEGs.csv"))

FastIIB_downregulated_KO_DEGs <- subset(FastIIB.de.markers, p_val_adj < 0.01 & avg_log2FC < -1)
# cat(rownames(FastIIB_downregulated_KO_DEGs), sep = "\n")
write.csv(FastIIB_downregulated_KO_DEGs, file.path(out_path, "/tables/FastIIB_downregulated_KO_DEGs.csv"))

FC_up_threshold <- 1
FC_down_threshold <- -1
df_up <- FastIIB_upregulated_KO_DEGs
df_down <- FastIIB_downregulated_KO_DEGs

# plot upregulated DEGs
Vol_up_DEG <- EnhancedVolcano(df_up,
  rownames(df_up),
  x = "avg_log2FC",
  y = "p_val_adj",
  pCutoff = significance_filter,
  FCcutoff = FC_up_threshold,
  pointSize = 2,
  drawConnectors = TRUE,
  title = "Fast IIB: Upregulated DEGs in KO vs. WT controls",
)
# ggsave(file.path(out_path, "/plots/FastIIB_KO_upregulated_DEGs.png"), plot = Vol_up_DEG, width = 10.1, height = 10.1)

# plot downregulated DEGs
Vol_down_DEG <- EnhancedVolcano(df_down,
  rownames(df_down),
  x = "avg_log2FC",
  y = "p_val_adj",
  pCutoff = significance_filter,
  FCcutoff = FC_down_threshold,
  pointSize = 2,
  drawConnectors = TRUE,
  title = "Fast IIB: Downregulated DEGs in KO vs. WT controls",
)
# ggsave(file.path(out_path, "/plots/FastIIB_KO_downregulated_DEGs.png"), plot = Vol_down_DEG, width = 10.1, height = 10.1)
```

```{r FastIIX all DEGs}
FastIIX.de.markers <- FindMarkers(seurat_obj, ident.1 = c("F7_KO_Fast IIX", "M4_KO_Fast IIX"), ident.2 = c("F15_WT_Fast IIX", "M11_WT_Fast IIX"), verbose = TRUE)
head(FastIIX.de.markers)
# write.csv(FastIIX.de.markers, file.path(out_path, "/tables/FastIIX_unfiltered_KO_DEGs.csv"))
df <- FastIIX.de.markers

# plot DEGs
significance_filter <- 0.01
FC_threshold <- 1
Vol_DEG <- EnhancedVolcano(df,
  rownames(df),
  x = "avg_log2FC",
  y = "p_val_adj",
  pCutoff = significance_filter,
  FCcutoff = FC_threshold,
  pointSize = 2,
  drawConnectors = TRUE,
  title = "Fast IIX: All DEGs in KO vs. WT controls",
)
# ggsave(file.path(out_path, "/plots/FastIIX_KO_DEGs.png"), plot = Vol_DEG, width = 10.1, height = 10.1)
```

```{r FastIIX subset up and down regulated DEGs}
# subset up and down regulated DEGs
FastIIX_upregulated_KO_DEGs <- subset(FastIIX.de.markers, p_val_adj < 0.01 & avg_log2FC > 1)
write.csv(FastIIX_upregulated_KO_DEGs, file.path(out_path, "/tables/FastIIX_upregulated_KO_DEGs.csv"))

FastIIX_downregulated_KO_DEGs <- subset(FastIIX.de.markers, p_val_adj < 0.01 & avg_log2FC < -1)
# write.csv(FastIIX_downregulated_KO_DEGs, file.path(out_path, "/tables/FastIIX_downregulated_KO_DEGs.csv"))

FC_up_threshold <- 1
FC_down_threshold <- -1
df_up <- FastIIX_upregulated_KO_DEGs
df_down <- FastIIX_downregulated_KO_DEGs

# plot upregulated DEGs
Vol_up_DEG <- EnhancedVolcano(df_up,
  rownames(df_up),
  x = "avg_log2FC",
  y = "p_val_adj",
  pCutoff = significance_filter,
  FCcutoff = FC_up_threshold,
  pointSize = 2,
  drawConnectors = TRUE,
  title = "Fast IIX: Upregulated DEGs in KO vs. WT controls",
)
# ggsave(file.path(out_path, "/plots/FastIIX_KO_upregulated_DEGs.png"), plot = Vol_up_DEG, width = 10.1, height = 10.1)

# plot downregulated DEGs
Vol_down_DEG <- EnhancedVolcano(df_down,
  rownames(df_down),
  x = "avg_log2FC",
  y = "p_val_adj",
  pCutoff = significance_filter,
  FCcutoff = FC_down_threshold,
  pointSize = 2,
  drawConnectors = TRUE,
  title = "Fast IIX: Downregulated DEGs in KO vs. WT controls",
)
# ggsave(file.path(out_path, "/plots/FastIIX_KO_downregulated_DEGs.png"), plot = Vol_down_DEG, width = 10.1, height = 10.1)
```

```{r FAPs all DEGs}
FAPs.de.markers <- FindMarkers(seurat_obj, ident.1 = c("F7_KO_FAPs", "M4_KO_FAPs"), ident.2 = c("F15_WT_FAPs", "M11_WT_FAPs"), verbose = TRUE)
head(FAPs.de.markers)
write.csv(FAPs.de.markers, file.path(out_path, "/tables/FAPs_unfiltered_KO_DEGs.csv"))
df <- FAPs.de.markers

# plot DEGs
significance_filter <- 0.01
FC_threshold <- 1
Vol_DEG <- EnhancedVolcano(df,
  rownames(df),
  x = "avg_log2FC",
  y = "p_val_adj",
  pCutoff = significance_filter,
  FCcutoff = FC_threshold,
  pointSize = 2,
  drawConnectors = TRUE,
  title = "FAPs: All DEGs in KO vs. WT controls",
)
ggsave(file.path(out_path, "/plots/FAPs_KO_DEGs.png"), plot = Vol_DEG, width = 10.1, height = 10.1)
```

```{r FAPs subset up and down regulated DEGs}
# subset up and down regulated DEGs
FAPs_upregulated_KO_DEGs <- subset(FAPs.de.markers, p_val_adj < 0.01 & avg_log2FC > 1)
write.csv(FAPs_upregulated_KO_DEGs, file.path(out_path, "/tables/FAPs_upregulated_KO_DEGs.csv"))

FAPs_downregulated_KO_DEGs <- subset(FAPs.de.markers, p_val_adj < 0.01 & avg_log2FC < -1)
write.csv(FAPs_downregulated_KO_DEGs, file.path(out_path, "/tables/FAPs_downregulated_KO_DEGs.csv"))

FC_up_threshold <- 1
FC_down_threshold <- -1
df_up <- FAPs_upregulated_KO_DEGs
df_down <- FAPs_downregulated_KO_DEGs

# plot upregulated DEGs
Vol_up_DEG <- EnhancedVolcano(df_up,
  rownames(df_up),
  x = "avg_log2FC",
  y = "p_val_adj",
  pCutoff = significance_filter,
  FCcutoff = FC_up_threshold,
  pointSize = 2,
  drawConnectors = TRUE,
  title = "FAPs: Upregulated DEGs in KO vs. WT controls",
)
ggsave(file.path(out_path, "/plots/FAPs_KO_upregulated_DEGs.png"), plot = Vol_up_DEG, width = 10.1, height = 10.1)

# plot downregulated DEGs
Vol_down_DEG <- EnhancedVolcano(df_down,
  rownames(df_down),
  x = "avg_log2FC",
  y = "p_val_adj",
  pCutoff = significance_filter,
  FCcutoff = FC_down_threshold,
  pointSize = 2,
  drawConnectors = TRUE,
  title = "FAPs: Downregulated DEGs in KO vs. WT controls",
)
ggsave(file.path(out_path, "/plots/FAPs_KO_downregulated_DEGs.png"), plot = Vol_down_DEG, width = 10.1, height = 10.1)
```

```{r Skeleton MuSc all DEGs}
Skeleton_MuSc.de.markers <- FindMarkers(seurat_obj, ident.1 = c("F7_KO_Skeleton MuSc", "M4_KO_Skeleton MuSc"), ident.2 = c("F15_WT_Skeleton MuSc", "M11_WT_Skeleton MuSc"), verbose = TRUE)
head(Skeleton_MuSc.de.markers)
write.csv(Skeleton_MuSc.de.markers, file.path(out_path, "/tables/MuSc_unfiltered_KO_DEGs.csv"))
df <- Skeleton_MuSc.de.markers

# plot DEGs
significance_filter <- 0.01
FC_threshold <- 1
Vol_DEG <- EnhancedVolcano(df,
  rownames(df),
  x = "avg_log2FC",
  y = "p_val_adj",
  pCutoff = significance_filter,
  FCcutoff = FC_threshold,
  pointSize = 2,
  drawConnectors = TRUE,
  title = "Skeleton Muscle Satellite Cells: All DEGs in KO vs. WT controls",
)
ggsave(file.path(out_path, "/plots/MuSc_KO_DEGs.png"), plot = Vol_DEG, width = 10.1, height = 10.1)
```

```{r Skeleton MuSc subset up and down regulated DEGs}
# subset up and down regulated DEGs
MuSc_upregulated_KO_DEGs <- subset(Skeleton_MuSc.de.markers, p_val_adj < 0.01 & avg_log2FC > 1)
write.csv(MuSc_upregulated_KO_DEGs, file.path(out_path, "/tables/MuSc_upregulated_KO_DEGs.csv"))

MuSc_downregulated_KO_DEGs <- subset(Skeleton_MuSc.de.markers, avg_log2FC < 0)
write.csv(MuSc_downregulated_KO_DEGs, file.path(out_path, "/tables/MuSc_downregulated_KO_DEGs.csv"))

FC_up_threshold <- 1
df_up <- MuSc_upregulated_KO_DEGs
df_down <- MuSc_downregulated_KO_DEGs

# plot upregulated DEGs
Vol_up_DEG <- EnhancedVolcano(df_up,
  rownames(df_up),
  x = "avg_log2FC",
  y = "p_val_adj",
  pCutoff = significance_filter,
  FCcutoff = FC_up_threshold,
  pointSize = 2,
  drawConnectors = TRUE,
  title = "Skeleton Muscle Satellite Cells: Upregulated DEGs in KO vs. WT controls",
)
ggsave(file.path(out_path, "/plots/MuSc_KO_upregulated_DEGs.png"), plot = Vol_up_DEG, width = 10.1, height = 10.1)

# plot downregulated DEGs
Vol_down_DEG <- EnhancedVolcano(df_down,
  rownames(df_down),
  x = "avg_log2FC",
  y = "p_val_adj",
  pCutoff = significance_filter,
  FCcutoff = 0,
  pointSize = 2,
  drawConnectors = TRUE,
  title = "Skeleton Muscle Satellite Cells: Downregulated DEGs in KO vs. WT controls",
)
ggsave(file.path(out_path, "/plots/MuSc_KO_downregulated_DEGs.png"), plot = Vol_down_DEG, width = 10.1, height = 10.1)
```

# Get DEGs in all cell types to check for ercc1 expression
```{r MTJ all DEGs}
MTJ.de.markers <- FindMarkers(seurat_obj, ident.1 = c("F7_KO_MTJ", "M4_KO_MTJ"), ident.2 = c("F15_WT_MTJ", "M11_WT_MTJ"), verbose = TRUE)
head(MTJ.de.markers)
write.csv(MTJ.de.markers, file.path(out_path, "/tables/MTJ_unfiltered_KO_DEGs.csv"))
```
```{r EC all DEGs}
EC.de.markers <- FindMarkers(seurat_obj, ident.1 = c("F7_KO_EC", "M4_KO_EC"), ident.2 = c("F15_WT_EC", "M11_WT_EC"), verbose = TRUE)
head(EC.de.markers)
write.csv(EC.de.markers, file.path(out_path, "/tables/EC_unfiltered_KO_DEGs.csv"))
```
```{r Pericyte all DEGs}
Pericyte.de.markers <- FindMarkers(seurat_obj, ident.1 = c("F7_KO_Pericyte", "M4_KO_Pericyte"), ident.2 = c("F15_WT_Pericyte", "M11_WT_Pericyte"), verbose = TRUE)
head(Pericyte.de.markers)
write.csv(Pericyte.de.markers, file.path(out_path, "/tables/Pericyte_unfiltered_KO_DEGs.csv"))
```
```{r Macrophages all DEGs}
Macrophages.de.markers <- FindMarkers(seurat_obj, ident.1 = c("F7_KO_Macrophages", "M4_KO_Macrophages"), ident.2 = c("F15_WT_Macrophages", "M11_WT_Macrophages"), verbose = TRUE)
head(Macrophages.de.markers)
write.csv(Macrophages.de.markers, file.path(out_path, "/tables/Macrophages_unfiltered_KO_DEGs.csv"))
```
```{r NMJ all DEGs}
NMJ.de.markers <- FindMarkers(seurat_obj, ident.1 = c("F7_KO_NMJ", "M4_KO_NMJ"), ident.2 = c("F15_WT_NMJ", "M11_WT_NMJ"), verbose = TRUE)
head(NMJ.de.markers)
write.csv(NMJ.de.markers, file.path(out_path, "/tables/NMJ_unfiltered_KO_DEGs.csv"))
```
```{r Tendon all DEGs}
Tendon.de.markers <- FindMarkers(seurat_obj, ident.1 = c("F7_KO_Tendon", "M4_KO_Tendon"), ident.2 = c("F15_WT_Tendon", "M11_WT_Tendon"), verbose = TRUE)
head(Tendon.de.markers)
write.csv(Tendon.de.markers, file.path(out_path, "/tables/Tendon_unfiltered_KO_DEGs.csv"))
```


# Save seurat_obj with the cell_id metadata column added for the major cell-type labels

```{r save final object}
saveRDS(cluster_seurat, file.path(out_path, "/objects/Seurat_cell_type_labeled_all_cells.rds"))
# reads the QC pre-processed seurat obj, which contains one layer for all sample cells (merged), and 10 cell-type labels
# celltype_seurat <- readRDS(file.path(out_path, "/objects/Seurat_cell_type_labeled_all_cells.rds"))
```
```{r inspect metadata}
print(head(celltype_seurat@meta.data))
```
```{r save anndata .h5ad}
# convert seurat object to anndata object
MuDataSeurat::WriteH5AD(celltype_seurat, file.path(out_path, "/objects/major_celltype_seurat.h5ad"), assay = "RNA")
```
