---
# DEG analysis script (M/F specific DEGs)

title: "scRNA analysis"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true # table of content true
    toc_depth: 1
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: false  ## if you want number sections at each table header
    theme: united 
    highlight: tango  
---

# Load libraries and Directories
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(stringr)
library(dplyr)
library(Seurat) # Use CRC's R lib
library(patchwork)
library(ggplot2) # version should be >=3.5.0 from CRAN
library(openxlsx)
library(harmony)
library(SingleR)
library(reticulate)
#library(MuDataSeurat) #to convert Rds to h5ad
library(EnhancedVolcano)

# Directories
out_path <- "/ix/djishnu/Akanksha/snRNA_TA_muscle_analysis/Seurat_analysis_outputs"
sample_names <- c("F7_KO", "F15_WT", "M4_KO", "M11_WT")
```

# Load the QC filtered seurat object with cell embeddings 
``` {r Reading seurat object}
# reads the QC pre-processed seurat obj, which contains one layer for all sample cells (merged), and 10 cell-type labels
seurat_obj <- readRDS(file.path(out_path, "/objects/Seurat_cell_type_labeled_all_cells.rds"))
print(head(seurat_obj@meta.data))
```
```{r print colors from a palette}
DiscretePalette(n = 4, palette = "alphabet", shuffle = FALSE)
```
```{r plot umaps}
umap_sample <- DimPlot(seurat_obj, reduction = "umap", label = TRUE, group.by = "orig.ident", cols = c("#FF5733", "#33FF57", "#3357FF", "#FFD700"), pt.size = 1)
ggsave(file.path(out_path, "/plots/mouse_sample_labels.png"), plot = umap_sample, width = 10.1, height = 10.1)
```
################################## UMAPs specific to gender (M/F) ##################################
```{r split the whole seurat object into gender specific objects}
# split the object based on condition metadata
gender_obj_list <- SplitObject(seurat_obj, split.by = "gender")
gender_obj_list$F
print(head(gender_obj_list$F@meta.data))
gender_obj_list$M
print(head(gender_obj_list$M@meta.data))
```
```{r plot MF sample umaps}
# plot umaps
# "#FF5733", "#33FF57" for females
umap_F_sample <- DimPlot(gender_obj_list$F, reduction = "umap", label = TRUE, group.by = "orig.ident", cols = c("#FF5733", "#33FF57"), pt.size = 1)
ggsave(file.path(out_path, "/plots/female_sample_labels.png"), plot = umap_F_sample, width = 10.1, height = 10.1)
umap_M_sample <- DimPlot(gender_obj_list$M, reduction = "umap", label = TRUE, group.by = "orig.ident", cols = c("#3357FF", "#FFD700"), pt.size = 1)
ggsave(file.path(out_path, "/plots/male_sample_labels.png"), plot = umap_M_sample, width = 10.1, height = 10.1)
```

# DEG analysis for Female and Male cell populations
################################## Filtering out non-nuclear DNA mapped genes before DEGs #######################
``` {r filtering mitochondrial mt-named genes}
# Get the gene names
all_genes <- rownames(seurat_obj)

# Filter out the genes that start with "mt-"
filtered_genes <- all_genes[!grepl("^mt-", all_genes)]

# Subset the Seurat object to exclude mitochondrial genes
seurat_obj <- subset(seurat_obj, features = filtered_genes)

# Check the new gene list without mt- genes
print(head(rownames(seurat_obj)))
```

################################## DEGs specific to gender (M/F) ##################################
```{r updating orig.ident with sample+condition_celltype for FindMarkers function}
metadata <- seurat_obj@meta.data
metadata$orig.ident <- paste(metadata$orig.ident, metadata$cell_type, sep = "_")
Idents(seurat_obj) <- metadata$orig.ident
seurat_obj@meta.data <- metadata
print(head(seurat_obj@meta.data))
```
```{r function for gender specific DEG}
DEG_analysis <- function(gender_ko, gender_wt, cell_type, gender) {
  # Define variables
  ko_ident <- paste0(gender_ko, "_", cell_type)
  wt_ident <- paste0(gender_wt, "_", cell_type)
  title <- paste0(cell_type, paste0(": DEGs in KO vs. WT controls for ", gender, " samples"))
  unfiltered_csv <- file.path(out_path, paste0("/tables/", gender, "_", cell_type, "_unfiltered_KO_DEGs_nmt.csv"))
  plot_png <- file.path(out_path, paste0("/plots/", gender, "_", cell_type, "_KO_DEGs_nmt.png"))
  upregulated_csv <- file.path(out_path, paste0("/tables/", gender, "_", cell_type, "_upregulated_KO_DEGs_nmt.csv"))
  downregulated_csv <- file.path(out_path, paste0("/tables/", gender, "_", cell_type, "_downregulated_KO_DEGs_nmt.csv"))

  # Find markers
  markers <- FindMarkers(seurat_obj, ident.1 = ko_ident, ident.2 = wt_ident, verbose = TRUE)
  write.csv(markers, unfiltered_csv)
  df <- markers

  # Plot DEGs
  significance_filter <- 0.01
  FC_threshold <- 1
  Vol_DEG <- EnhancedVolcano(df,
    rownames(df),
    x = "avg_log2FC",
    y = "p_val_adj",
    pCutoff = significance_filter,
    FCcutoff = FC_threshold,
    pointSize = 2,
    drawConnectors = TRUE,
    title = title
  )
  ggsave(plot_png, plot = Vol_DEG, width = 10.1, height = 10.1)

  # Subset up and downregulated DEGs
  upregulated_KO_DEGs <- subset(markers, p_val_adj < 0.01 & avg_log2FC > 1)
  write.csv(upregulated_KO_DEGs, upregulated_csv)

  downregulated_KO_DEGs <- subset(markers, p_val_adj < 0.01 & avg_log2FC < -1)
  write.csv(downregulated_KO_DEGs, downregulated_csv)
}
# run the function: params - sample_KO, sample_WT, cell_type (Fast IIX, Fast IIB, FAPs, Skeleton MuSc), Female/Male
DEG_analysis("F7_KO", "F15_WT", "Fast IIX", "Female")
```
################################## Genes of interest in Cell embeddings ##################################
# Genes of interest: CDKN1A/MYH8/COL19A1/LRRK2/EDAR2/ and ATP2A1; common genes between MF (MALAT1, mt-co2, mt-co3, mt-atp6)
```{r ercc1, ercc4}
FeaturePlot(seurat_obj, features = c("Ercc1", "Ercc4"), min.cutoff = "q10", max.cutoff = "q90", ncol = 2)
ggplot2::ggsave(file.path(out_path, "/plots/ercc1_ercc4.png"), width = 10.1, height = 10.1)
```
```{r tpm2}
FeaturePlot(seurat_obj, features = c("Xist", "Ptprg"), min.cutoff = "q10", max.cutoff = "q90", ncol = 1)
ggplot2::ggsave(file.path(out_path, "/plots/Xist_Ptprg_all_cells.png"), width = 10.1, height = 10.1)
```
```{r feature plot for common genes in MF}
FeaturePlot(seurat_obj, features = c("Malat1", "mt-Co2", "mt-Co3", "mt-Atp6"), min.cutoff = "q10", max.cutoff = "q90", ncol = 4)
ggplot2::ggsave(file.path(out_path, paste0("/plots/common_MF_DEGs.png")), width = 12, height = 6, dpi = 300)
```
```{r feature plot for genes of interest (per-gene)}
FeaturePlot(seurat_obj, features = c("Cdkn1a", "Myh8", "Col19a1", "Lrrk2", "Eda2r", "Atp2a1"), min.cutoff = "q10", max.cutoff = "q90", ncol = 3)
ggplot2::ggsave(file.path(out_path, paste0("/plots/genes_of_interest_cell_embeddings.png")), width = 12, height = 6, dpi = 300)
```
```{r feature plot for genes of interest (gene-combinations)}
# FeaturePlot(seurat_obj, features = c("Lrrk2", "Eda2r"), min.cutoff = "q10", max.cutoff = "q90", blend = TRUE)
# ggplot2::ggsave(file.path(out_path, paste0("/plots/coexpression_Lrrk2_Eda2r.png")), width = 12, height = 6, dpi = 300)
```
```{r split the whole seurat object into condition specific objects}
# split the object based on condition metadata
condition_obj_list <- SplitObject(seurat_obj, split.by = "condition")
condition_obj_list$KO
print(head(condition_obj_list$KO@meta.data))
condition_obj_list$WT
print(head(condition_obj_list$WT@meta.data))
```
```{r function for generating feature plots}
# Function to generate feature plots for a specific object type
plot_feature_for_object <- function(condition_obj_list, plot_str, obj_type, features, out_path) {
  # Generate the feature plot
  p <- FeaturePlot(condition_obj_list[[obj_type]],
    features = features,
    min.cutoff = "q10",
    max.cutoff = "q90",
    ncol = 1
  )

  # Save the plot
  ggsave(file.path(out_path, paste0("/plots/", plot_str, obj_type, "_cells.png")),
    plot = p,
    width = 12,
    height = 6,
    dpi = 300
  )
}
# Plot for KO/WT/all cells
# common_MF_DEGs <- c("Malat1", "mt-Co2", "mt-Co3", "mt-Atp6")
# plot_feature_for_object(condition_obj_list, "common_MF_DEGs_", "WT", common_MF_DEGs, out_path)
genes_of_interest <- c("Tpm2")
plot_feature_for_object(condition_obj_list, "tpm2_", "WT", genes_of_interest, out_path)
# dna_damage_genes <- c("Ercc1", "Ercc4")
# plot_feature_for_object(condition_obj_list, "dna_damage_genes_", "KO", dna_damage_genes, out_path)
```

################################## what is unique to a specific cell population? ##################################
```{r hover plot for specific cell population study in all samples}
type1_plot <- DimPlot(seurat_obj, reduction = "umap", label = TRUE, group.by = "orig.ident", pt.size = 1)
select.cells <- CellSelector(plot = type1_plot)
head(select.cells)
```
```{r finding markers for the selected cells}
Idents(seurat_obj, cells = select.cells) <- "NewCells"

# Now, we find markers that are specific to the new cells, and find clear DC markers
newcells.markers <- FindMarkers(seurat_obj,
  ident.1 = "NewCells", ident.2 = "M11_WT", min.diff.pct = 0.3
)
head(newcells.markers)
write.csv(newcells.markers, file.path(out_path, paste0("/tables/newcells_vs_MWT_markers.csv")))
df <- newcells.markers
# Plot DEGs
significance_filter <- 0.01
FC_threshold <- 1
Vol_DEG <- EnhancedVolcano(df,
  rownames(df),
  x = "avg_log2FC",
  y = "p_val_adj",
  pCutoff = significance_filter,
  FCcutoff = FC_threshold,
  pointSize = 2,
  drawConnectors = TRUE,
  title = "NewCells: DEGs as compared with M_WT cells"
)
ggsave(file.path(out_path, paste0("/plots/newcells_vs_MWT_DEGs.png")), plot = Vol_DEG, width = 10.1, height = 10.1)
```
################################## subsetting to only four celltypes for heatmap ##################################
```{r separate selected celltypes into a different object}
# separate the selected cell types into a new object
muscle_obj <- subset(seurat_obj, idents = c("Fast IIB", "Fast IIX", "FAPs", "Skeleton MuSc"))
print(table(Idents(muscle_obj)))
print(head(muscle_obj@meta.data))
```
```{r split the whole seurat object into condition specific objects}
# split the object based on condition metadata
condition_muscle_obj_list <- SplitObject(muscle_obj, split.by = "condition")
condition_muscle_obj_list$KO
print(head(condition_muscle_obj_list$KO@meta.data))
condition_muscle_obj_list$WT
print(head(condition_muscle_obj_list$WT@meta.data))
```

```
```{r function for generating heatmaps}
# Function to generate feature plots for a specific object type
plot_heat_for_object <- function(object, obj_type, features, out_path) {
  # scale data for these features
  object <- ScaleData(object, features = features)
  # plot heatmap
  p <- DoHeatmap(
    object = object, features = features,
    group.by = "ident"
  )
  # Save the plot
  ggsave(file.path(out_path, paste0("/plots/heat_interest_", obj_type, "_cells.png")),
    plot = p,
    width = 12,
    height = 6,
    dpi = 300
  )
}
heatmap_markers <- c("Cdkn1a", "Myh8", "Col19a1", "Lrrk2", "Eda2r", "Atp2a1")
# Plot for KO cells
plot_heat_for_object(condition_muscle_obj_list$WT, "WT", heatmap_markers, out_path)
```

```{r correlation of GE plot}
correl <- FeatureScatter(object = muscle_obj, feature1 = "Cdkn1a", feature2 = "Myh8")
ggsave(file.path(out_path, "/plots/correl_cdkn1a_myh8.png"), plot = correl, width = 10.1, height = 10.1)
```